---
title: "fear-contagion-ami_data-wrangling"
author: "Jack L Manera"
format: html
---


# 1. Setup environment

*Install & Load Required Packages*
```{r packages}
#If 'pacman' is not already installed please un-hash and run the line below
#install.packages("pacman")
#remotes::install_github("JLManera/cleanRfish")

# Core packages for manipulation and plotting
pacman::p_load(tidyverse, ggplot2, zoo, readxl, cleanRfish, stringr, moveHMM, patchwork, future, furrr, sf, changepoint, install = TRUE, update = TRUE)
```

# 2. Clean raw tracks

## 2.1 Locate trajectory files

Traverse the idtracker.ai output folder structure and recursively find all trajectories.csv files (i.e., raw tracks).

```{r paths, message=FALSE, error=FALSE}
idtracker_raw_output_directory <- "./2-data/1-raw_tracks"

csv_files <- list.files(
  path = idtracker_raw_output_directory,
  pattern = "trajectories\\.csv$",
  recursive = TRUE,
  full.names = TRUE
) |>
  (\(x) x[!grepl("pretrained", x)])() #removing data from the pretrained model (this was used as a starting set of weights for the tracking model)

# focal tracks
focal_csvs <- csv_files |> 
  keep(~ str_detect(.x, regex("focal", ignore_case = TRUE)))

# demonstrator tracks
demonstrators_csvs <- csv_files |> 
  keep(~ str_detect(.x, regex("demonstrator", ignore_case = TRUE)))
```


## 2.2 Read in assisting dataframes

This reads in both calibration data needed to convert pixels to cm's and infomrmation regarding track reconstruction.
```{r calibration}
#this is a csv that indicates whether track reconstruction was better of worse than the raw track itself and thus is used to decide whether a track is reconstructed or not.
reconstruct_or_not_df <- read_csv("./2-data/video_validation.csv") |>
  mutate(session = str_remove(Video, "\\..*$"))

#this is the video pixel calibration to convert pixels to cms
distance_calibration_df <- read_csv("./2-data/pixel_distances.csv") |> 
  select(c(id, distance_px)) |>
  mutate(
    day_trial_camera = str_extract(id, "^d\\d+-t\\d+-c\\d+")
  )

# Calculate average pixels per day-trial-camera grouping
group_avg_pixels <- distance_calibration_df |>
  group_by(day_trial_camera) |>
  summarize(
    avg_distance_px = mean(distance_px, na.rm = TRUE),
    n_measurements = n(),
    .groups = "drop"
  ) |>
  mutate(pixels_per_cm = avg_distance_px / 45,
         cam = substr(day_trial_camera, 7, 10)) 

  
ggplot(group_avg_pixels, aes(x = pixels_per_cm)) +
  geom_histogram(bins = 30) +
  labs(title = "Calibration Distribution (averaged per day-trial-camera)", 
       x = "Pixels per cm", y = "Count") +
  facet_wrap(~cam)
```

## 2.3 Metadata extraction

```{r}
extract_meta_information <- function(path) {
  # Extract the session folder name
  session_name <- stringr::str_extract(path, "session_[^/]+") |>
    stringr::str_remove("^session_")

  # Extract day, trial, camera numbers
  day <- stringr::str_match(session_name, "d(\\d+)")[, 2]
  trial <- stringr::str_match(session_name, "t(\\d+)")[, 2]
  camera <- stringr::str_match(session_name, "c(\\d+)")[, 2]

  # Phase: baseline if contains "-b-", else trial if contains "-t-"
  phase <- case_when(
    stringr::str_detect(session_name, "-b-") ~ "baseline",
    stringr::str_detect(session_name, "-t-") ~ "trial",
    TRUE ~ NA_character_
  )

  # ID: whatever comes after the last "-b-" or "-t-"
  ID <- stringr::str_extract(session_name, "(?<=-[bt]-).+$")
  
  treatment <- substr(ID, 2, 2)
  
  information <- ifelse(substr(ID, 5,5) == "d", "personal", ifelse(substr(ID, 5,5) == "i", "social", "error"))

  tibble::tibble(
    path = path,
    session_name = session_name,
    day = as.integer(day),
    trial = as.integer(trial),
    camera = as.integer(camera),
    phase = phase,
    ID = ID,
    treatment = treatment,
    information = information
  )
}
```

## 2.4 Main processing pipeline

This section loops through every tracked session and performs the core data-processing steps. Uses parallel processing to speed up computation across multiple CPU cores.

```{r process_save_each_trial}
output_track_dir <- "./2-data/2-smoothed_tracks"

# Set up parallel processing - uses all available cores minus 1
plan(multisession, workers = max(1, availableCores() - 1))

# Define the processing function for a single session
process_single_session <- function(csv_file, demonstrators_csvs, group_avg_pixels, reconstruct_or_not_df) {
  
  meta_info  <- extract_meta_information(csv_file)
  session_id <- meta_info$session_name[1]
  
  #wrap entire processing in tryCatch to identify if session fails
  tryCatch({
    
    trajectories_df <- readr::read_csv(csv_file, show_col_types = FALSE) |> 
      rename(x1 = trajectories1,
             y1 = trajectories2)
    
    reconstruct_or_not <- reconstruct_or_not_df |> 
      filter(session == session_id)
    
    # Handle case where reconstruct_or_not is empty (no validation data for this session)
    if (nrow(reconstruct_or_not) == 0) {
      warning(sprintf("No validation data found for %s. Using reconstructed tracks by default.", session_id))
      focal_use_reconstructed <- TRUE
      dem_use_reconstructed <- TRUE
    } else {
      focal_use_reconstructed <- reconstruct_or_not$focal_fixed_with_cleanRfish[1] == "y"
      dem_use_reconstructed <- reconstruct_or_not$dem_fixed_with_cleanRfish[1] == "y"
    }
    
    demonstrator_csv_file <- demonstrators_csvs[grepl(paste0(sub("-focal$", "-demonstrator", session_id)), demonstrators_csvs)]
    
    # Track the number of demonstrators found
    n_demonstrators <- 0
    
    if (length(demonstrator_csv_file) == 1) {
      demonstrator_df <- readr::read_csv(demonstrator_csv_file, show_col_types = FALSE)
      
      # Dynamically rename demonstrator columns based on what exists
      # The demonstrator file may have 1, 2, or 3 fish (x1/y1, x2/y2, x3/y3)
      rename_map <- c()
      if ("x1" %in% names(demonstrator_df)) { rename_map <- c(rename_map, x2 = "x1", y2 = "y1"); n_demonstrators <- n_demonstrators + 1 }
      if ("x2" %in% names(demonstrator_df)) { rename_map <- c(rename_map, x3 = "x2", y3 = "y2"); n_demonstrators <- n_demonstrators + 1 }
      if ("x3" %in% names(demonstrator_df)) { rename_map <- c(rename_map, x4 = "x3", y4 = "y3"); n_demonstrators <- n_demonstrators + 1 }
      
      if (length(rename_map) > 0) {
        demonstrator_df <- demonstrator_df |> rename(!!!rename_map)
      }
      
      # Combine focal and demonstrator data
      trajectories_df <- merge(trajectories_df, demonstrator_df, by = "time")
      } else {
      warning(sprintf("No matching demonstrator file found for %s.", session_id))
    }
    
    #trim the baseline recordings so that they are all of equal length
    if(stringr::str_detect(session_id, "-b-")) {
      trajectories_df <- trajectories_df |> 
        filter(time >= 300 & time <= 1500) |> 
        mutate(time = time - 300)
      }
    
    #reconstruct using cleanRfish
    reconstructed_path <- tryCatch(
      cleanRfish::clean_track(
        trajectories_df,
        z_thresh = 5,
        window = 20,
        speed_threshold_quantile = 1,
        smooth_method = "savitzky_golay",
        min_movement = 500,
        diagnostic_plots = FALSE,
        min_candidates = 5
      ),
      error = function(e) {
        warning(sprintf(
          "Trajectory reconstruction failed for %s: %s\n.",
          session_id,
          e$message
        ))
        return(list(tracks = trajectories_df, error = TRUE))
      }
    )
    
    # Update the tracks dataframe: remove existing x_smooth/y_smooth and recreate based on validation
    reconstructed_path$tracks <- reconstructed_path$tracks |>
      mutate(
        # For focal individual (individual_number == 1):
        # If reconstruction was NOT helpful ("n"), use original; otherwise use reconstructed
        x_smooth = case_when(
          individual_number == 1 & !focal_use_reconstructed ~ x_original,
          individual_number == 1 & focal_use_reconstructed ~ x_reconstructed,
          # For demonstrator individuals (2, 3, 4):
          individual_number %in% c(2, 3, 4) & !dem_use_reconstructed ~ x_original,
          individual_number %in% c(2, 3, 4) & dem_use_reconstructed ~ x_reconstructed,
          TRUE ~ x_smooth
        ),
        y_smooth = case_when(
          individual_number == 1 & !focal_use_reconstructed ~ y_original,
          individual_number == 1 & focal_use_reconstructed ~ y_reconstructed,
          individual_number %in% c(2, 3, 4) & !dem_use_reconstructed ~ y_original,
          individual_number %in% c(2, 3, 4) & dem_use_reconstructed ~ y_reconstructed,
          TRUE ~ y_smooth
        )
      ) |> 
      group_by(individual_number) |> 
      arrange(time) |> 
      mutate(x_filled = zoo::na.approx(x_smooth, na.rm = FALSE),
             y_filled = zoo::na.approx(y_smooth, na.rm = FALSE)) |>
      ungroup()
  
  # Helper function to apply sgolayfilt 
  safe_sgolay <- function(x, p, n) {
    if (length(x) >= n) {
      return(signal::sgolayfilt(x, p = p, n = n))
    } else {
      return(x)
    }
  }
  
  # Apply Savitzky-Golay smoothing 
  smoothed_subset <- reconstructed_path$tracks |>
    filter(!is.na(x_filled), !is.na(y_filled)) |>
    group_by(individual_number) |>
    arrange(time) |>
    mutate(
      x_smooth_sg = safe_sgolay(x_filled, p = 3, n = 13),
      y_smooth_sg = safe_sgolay(y_filled, p = 3, n = 13)
    ) |>
    ungroup() |>
    select(time, individual_number, x_smooth_sg, y_smooth_sg)
  
  # Join smoothed values back to preserve temporal alignment
  reconstructed_path$tracks <- reconstructed_path$tracks |>
    left_join(smoothed_subset, by = c("time", "individual_number")) |>
    mutate(
      x_smooth = x_smooth_sg,
      y_smooth = y_smooth_sg
    ) |>
    select(-x_smooth_sg, -y_smooth_sg, -x_filled, -y_filled)
  
  #attach metadata and apply calibration
  linking_key <- paste0("d", meta_info$day, "-t", meta_info$trial, "-c", meta_info$camera)
  calib_row <- group_avg_pixels |> filter(day_trial_camera == linking_key)
  
  if (nrow(calib_row) == 0) {
    warning(sprintf("No calibration data found for %s (key: %s). Using NA.", session_id, linking_key))
    px_per_cm <- NA_real_
  } else {
    px_per_cm <- calib_row$pixels_per_cm[1]
  }
  
  smooth_df <- reconstructed_path$tracks |>
    mutate(
      path = meta_info$path,
      session_name = meta_info$session_name,
      day = meta_info$day,
      trial = meta_info$trial,
      camera = meta_info$camera,
      phase = meta_info$phase,
      ID = meta_info$ID,
      treatment = meta_info$treatment,
      information = meta_info$information,
      pixels_per_cm = px_per_cm
    ) |>
    mutate(
      x_smooth = x_smooth / px_per_cm,
      y_smooth = y_smooth / px_per_cm
    ) |>
    group_by(individual_number) |>
    arrange(time) |>
    mutate(
      velocity = sqrt((x_smooth - lag(x_smooth))^2 + (y_smooth - lag(y_smooth))^2) / (time - lag(time)),
      acceleration = (velocity - lag(velocity)) / (time - lag(time)),
      heading = atan2(y_smooth - lag(y_smooth), x_smooth - lag(x_smooth)),
      turn_angle = heading - lag(heading),
      turn_angle2 = atan2(sin(turn_angle), cos(turn_angle))
    ) |>
    ungroup()
  
  wide_df <- smooth_df |>
    select(-segment, -x_reconstructed, -y_reconstructed) |>
    tidyr::pivot_wider(
      id_cols = c(time, path, session_name, day, trial, camera, phase, ID, treatment, information, pixels_per_cm),
      names_from = individual_number,
      values_from = c(x_original, y_original, x_smooth, y_smooth, velocity, acceleration, turn_angle, turn_angle2, heading),
      names_sep = "_ind"
    )
  
  out_path <- file.path(output_track_dir, paste0(session_id, ".csv"))
  write_csv(wide_df, out_path)
  
  return(list(session_id = session_id, success = TRUE, n_demonstrators = n_demonstrators))
  }, error = function(e) {
    #this will catch any error and report which session failed
    warning(sprintf("failed processing session %s: %s", session_id, e$message))
    return(list(session_id = session_id, success = FALSE, n_demonstrators = 0, 
                error_msg = e$message))
  })
}

# Process all sessions in parallel with progress updates
results <- future_map(
  focal_csvs,
  ~ process_single_session(.x, demonstrators_csvs, group_avg_pixels, reconstruct_or_not_df),
  .options = furrr_options(seed = TRUE),
  .progress = TRUE
)

# Reset to sequential processing
plan(sequential)
```


# 3. HMM Behavioural states

## 3.1 read in smoothed tracks
```{r load_smoothed_tracks, message=FALSE}
smoothed_track_dir <- "./2-data/2-smoothed_tracks"

smoothed_files <- list.files(
  path = smoothed_track_dir,
  pattern = "\\.csv$",
  full.names = TRUE) |>
  (\(x) x[!grepl("^\\._", basename(x))])()

combined_df <- map_dfr(smoothed_files, read_csv, show_col_types = FALSE)
```

## 3.2 Prepare all individuals in long format

Pivot all individuals (focal + demonstrators) into one long-format dataframe. `moveHMM::prepData()` uses the `ID` column to treat each individual-session as an independent track, so we fit a single HMM pooling all data.

```{r prepare_hmm_long, message=FALSE, warning=FALSE}
# Pivot all individuals to long format and generate the columns required for moveHMM
hmm_long <- combined_df |>
  select(time, session_name, matches("^[xy]_smooth_ind[1-4]$")) |>
  pivot_longer(
    cols = matches("^[xy]_smooth_ind[1-4]$"),
    names_to = c(".value", "ind"),
    names_pattern = "^(x_smooth|y_smooth)_(ind[1-4])$"
  ) |>
  rename(x = x_smooth, y = y_smooth) |>
  filter(!is.na(x) & !is.na(y)) |>
  mutate(ID = paste0(session_name, "__", ind)) |>
  arrange(ID, time) |>
  # Average to 1-second resolution as the HMM starts to fail at higher resolutions
  mutate(second = floor(time)) |>
  group_by(ID, session_name, ind, second) |>
  summarise(x = mean(x), y = mean(y), .groups = "drop") |>
  rename(time = second)

cat(sprintf(
  "Long-format HMM data: %d rows, %d unique tracks (%d sessions × up to 4 individuals)\n",
  nrow(hmm_long), n_distinct(hmm_long$ID), n_distinct(hmm_long$session_name)
))


data_hmm <- prepData(hmm_long, type = "UTM", coordNames = c("x", "y"))
```

## 3.3 Fit single global HMM

```{r fit_global_hmm}
mu0 <- c(0.01, 1)
sigma0 <- c(0.1, 1)
zeromass0 <- c(0.05, 0.00001)
stepPar0 <- c(mu0, sigma0, zeromass0)

angleMean0 <- c(0, 0)
kappa0 <- c(0.1, 1)
anglePar0 <- c(angleMean0, kappa0)

m_all <- fitHMM(
  data = data_hmm,
  nbStates = 2,
  stepDist = "gamma",
  stepPar0 = stepPar0,
  angleDist = "vm",
  anglePar0 = anglePar0
)

m_all
```

## 3.4 Decode states and save updated CSVs

```{r decode_and_save}
# Viterbi decode all tracks at once
states <- viterbi(m_all)
freezing_state <- which.min(m_all$mle$stepPar[1, ])

hmm_decoded <- data_hmm |>
  as_tibble() |>
  mutate(
    state = states,
    behaviour = if_else(state == freezing_state, "freezing", "moving")
  )

cat(sprintf("Decoded %d observations. Freezing state = %d\n", nrow(hmm_decoded), freezing_state))

behaviour_lookup <- hmm_decoded |>
  select(ID, time, behaviour) |>
  tidyr::separate_wider_delim(ID, delim = "__", names = c("session_name", "ind")) |>
  rename(time_second = time)

# Pivot to wide so each individual gets its own behaviour column
behaviour_wide <- behaviour_lookup |>
  mutate(ind = paste0("behaviour_", ind)) |>
  pivot_wider(names_from = ind, values_from = behaviour)

# Write updated CSVs
smoothed_track_dir <- "./2-data/2-smoothed_tracks"
output_track_dir_updated <- "./2-data/3-behaviour_tracks"

all_sessions <- unique(behaviour_wide$session_name)
updated_count <- 0

for (session_id in all_sessions) {
  csv_path <- file.path(smoothed_track_dir, paste0(session_id, ".csv"))
  if (!file.exists(csv_path)) {
    warning(sprintf("CSV not found for session %s", session_id))
    next
  }

  session_df <- read_csv(csv_path, show_col_types = FALSE) |>
    select(-any_of(c("behaviour", paste0("behaviour_ind", 1:4)))) |>
    mutate(time_second = floor(time))

  session_beh <- behaviour_wide |>
    filter(session_name == session_id) |>
    select(-session_name)

  session_df <- session_df |>
    left_join(session_beh, by = "time_second") |>
    select(-time_second)

  write_csv(session_df, file.path(output_track_dir_updated, paste0(session_id, ".csv")))
  updated_count <- updated_count + 1
  if (updated_count %% 25 == 0) cat("Updated", updated_count, "files...\n")
}

cat(sprintf("\nUpdated %d CSV files with behaviour columns (behaviour_ind1–ind4).\n", updated_count))
```

## 3.5 Plot a sample of sessions with behavioural states

```{r plot_sample_sessions}
# Keep only focal fish (ind1) for the sample plot
hmm_focal_df_with_states <- hmm_decoded |>
  filter(grepl("__ind1$", ID))

sample_sessions <- unique(hmm_focal_df_with_states$ID)[1:min(14, length(unique(hmm_focal_df_with_states$ID)))]

df_sample <- hmm_focal_df_with_states |>
  filter(ID %in% sample_sessions)

ggplot(df_sample, aes(x = time, y = step, color = behaviour)) +
  geom_point(alpha = 0.6, size = 1) +
  scale_color_manual(
    values = c("freezing" = "#E31A1C", "moving" = "#1F78B4"),
    labels = c("Freezing", "Moving")
  ) +
  labs(
    title = "Movement Speed with HMM-Detected Behavioural States (Sample Sessions)",
    x = "Time (s)",
    y = "Step Length (cm/s)",
    color = "Behavioural State"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    plot.title = element_text(face = "bold"),
    strip.text = element_text(size = 7)
  ) +
  facet_wrap(~ID, ncol = 2)
```

# 4. Spatial use

Here we go and collect the ROI lists from the corresponding .toml files for each "focal" track to get the coordinates of the arena (provided in pixels). Then I use 'sf' to get distance from the edge of the arena (thigmotaxis) and the distance from the top of the arena (i.e., the side closest to the demonstrators; social proximity).

Specifically, this code loops through each smoothed track .csv and find its corresponding .toml file, reads in the ROI coordinates, create an sf object with the ROI, and then calculates the distance from each smoothed point to the edge of the ROI and the top of the ROI (the demonstrators side). This code also generates a plot for each track showing the fishes movement inside the ROI for verification of both the tracking quality and the specification of the arena and demonstrator side.

## 4.1 Functions to extract arena

```{r spatial_setup}
# Function to parse ROI polygon from TOML format
parse_roi_polygon <- function(roi_string) {
  coords_match <- str_extract_all(roi_string, "\\[\\d+\\.?\\d*, \\d+\\.?\\d*\\]")[[1]]
  
  coords <- lapply(coords_match, function(coord) {
    nums <- as.numeric(str_extract_all(coord, "\\d+\\.?\\d*")[[1]])
    return(nums)
  })
  
  coord_matrix <- do.call(rbind, coords)
  coord_matrix <- rbind(coord_matrix, coord_matrix[1, ])
  
  return(coord_matrix)
}

# Function to read TOML file and extract ROI
read_toml_roi <- function(toml_path) {
  if (!file.exists(toml_path)) {
    warning(sprintf("TOML file not found: %s", toml_path))
    return(NULL)
  }

  lines <- readLines(toml_path)
  
  roi_lines <- lines[grepl("roi_list", lines) | grepl("Polygon", lines)]
  
  positive_roi_lines <- roi_lines[grepl("\\+\\s*Polygon", roi_lines)]
  
  if (length(positive_roi_lines) == 0) {
    positive_roi_lines <- roi_lines[!grepl("-\\s*Polygon", roi_lines) & grepl("Polygon", roi_lines)]
  }
  
  roi_string <- paste(positive_roi_lines, collapse = " ")
  
  if (grepl("Polygon", roi_string)) {
    coords <- parse_roi_polygon(roi_string)
    return(coords)
  } else {
    warning(sprintf("No positive polygon found in: %s", toml_path))
    return(NULL)
  }
}

# Function to find the "top edge" of an irregular polygon (demonstrator side) [surprisingly more complex to do than I thought it would be]
# Approach: 
# 1. Get the bounding box to find where "top" is
# 2. Find all polygon edges near the top (within 15% of bbox top)
# 3. Select the longest edge - that's the main demonstrator side
# 4. Add any connected edges that share a vertex AND are >= 40% of best_edge length
get_demonstrator_edge <- function(polygon_coords, arena_sf) {
  bbox <- st_bbox(arena_sf)
  arena_height <- bbox["ymax"] - bbox["ymin"]
  top_threshold <- bbox["ymin"] + (arena_height * 0.15)
  
  n_points <- nrow(polygon_coords) - 1
  coords <- polygon_coords[1:n_points, ]
  
  edges <- data.frame(
    idx = 1:n_points,
    x1 = coords[, 1],
    y1 = coords[, 2],
    x2 = c(coords[-1, 1], coords[1, 1]),
    y2 = c(coords[-1, 2], coords[1, 2])
  )
  
  edges$mid_y <- (edges$y1 + edges$y2) / 2
  edges$length <- sqrt((edges$x2 - edges$x1)^2 + (edges$y2 - edges$y1)^2)
  
  top_edges <- edges[edges$mid_y <= top_threshold, ]
  
  if (nrow(top_edges) == 0) {
    best_edge <- edges[which.min(edges$mid_y), ]
  } else {
    best_edge <- top_edges[which.max(top_edges$length), ]
  }
  
  length_threshold <- best_edge$length * 0.4
  best_edge_angle <- atan2(best_edge$y2 - best_edge$y1, best_edge$x2 - best_edge$x1) * 180 / pi
  
  # Helper function to check if two edges share a vertex
  shares_vertex <- function(e1, e2) {
    (e2$x1 == e1$x1 && e2$y1 == e1$y1) ||
    (e2$x1 == e1$x2 && e2$y1 == e1$y2) ||
    (e2$x2 == e1$x1 && e2$y2 == e1$y1) ||
    (e2$x2 == e1$x2 && e2$y2 == e1$y2)
  }
  
  # Helper function to check if two edges are nearly parallel (within 10 degrees)
  is_nearly_parallel <- function(e1_angle, e2) {
    e2_angle <- atan2(e2$y2 - e2$y1, e2$x2 - e2$x1) * 180 / pi
    angle_diff <- abs(e1_angle - e2_angle)
    angle_diff <- min(angle_diff, 360 - angle_diff)
    if (angle_diff > 90) angle_diff <- 180 - angle_diff
    return(angle_diff <= 10)
  }
  
  connected_edges <- edges[edges$idx != best_edge$idx, ]
  connected_edges <- connected_edges[sapply(1:nrow(connected_edges), function(i) {
    shares_vertex(best_edge, connected_edges[i, ])
  }), ]
  
  qualifying_connected <- connected_edges[
    connected_edges$length >= length_threshold & 
    sapply(1:nrow(connected_edges), function(i) {
      is_nearly_parallel(best_edge_angle, connected_edges[i, ])
    }), ]
  
  all_selected_edges <- rbind(best_edge, qualifying_connected)
  
  all_points <- unique(rbind(
    data.frame(x = all_selected_edges$x1, y = all_selected_edges$y1),
    data.frame(x = all_selected_edges$x2, y = all_selected_edges$y2)
  ))
  
 if (nrow(all_points) == 2) {
    top_edge <- matrix(
      c(best_edge$x1, best_edge$y1,
        best_edge$x2, best_edge$y2),
      ncol = 2, byrow = TRUE
    )
    return(st_linestring(top_edge))
  }
  
  edge_list <- lapply(1:nrow(all_selected_edges), function(i) {
    list(
      p1 = c(all_selected_edges$x1[i], all_selected_edges$y1[i]),
      p2 = c(all_selected_edges$x2[i], all_selected_edges$y2[i])
    )
  })
  
  point_counts <- table(paste(
    c(all_selected_edges$x1, all_selected_edges$x2),
    c(all_selected_edges$y1, all_selected_edges$y2),
    sep = ","
  ))
  endpoints <- names(point_counts[point_counts == 1])
  
  if (length(endpoints) >= 1) {
    start_coords <- as.numeric(strsplit(endpoints[1], ",")[[1]])
    
    ordered_points <- matrix(start_coords, ncol = 2, byrow = TRUE)
    current_point <- start_coords
    used_edges <- logical(length(edge_list))
    
    for (iter in 1:length(edge_list)) {
      found <- FALSE
      for (i in which(!used_edges)) {
        edge <- edge_list[[i]]
        if (all(edge$p1 == current_point)) {
          ordered_points <- rbind(ordered_points, edge$p2)
          current_point <- edge$p2
          used_edges[i] <- TRUE
          found <- TRUE
          break
        } else if (all(edge$p2 == current_point)) {
          ordered_points <- rbind(ordered_points, edge$p1)
          current_point <- edge$p1
          used_edges[i] <- TRUE
          found <- TRUE
          break
        }
      }
      if (!found) break
    }
    
    return(st_linestring(ordered_points))
  } else {
    top_edge <- matrix(
      c(best_edge$x1, best_edge$y1,
        best_edge$x2, best_edge$y2),
      ncol = 2, byrow = TRUE
    )
    return(st_linestring(top_edge))
  }
}
```

## 4.2 Process each session - calculate spatial metrics

```{r spatial_processing, message=FALSE, warning=FALSE}
smoothed_track_dir <- "./2-data/3-behaviour_tracks"
toml_dir <- "./2-data/1-raw_tracks"
spatial_output_dir <- "./3-figs/spatial_checks"

smoothed_files <- list.files(
  path = smoothed_track_dir,
  pattern = "\\.csv$",
  full.names = TRUE) |>
  (\(x) x[!grepl("^\\._", basename(x))])() |>
  (\(x) x[grepl("focal", x)])() 

spatial_results <- list()

for (i in seq_along(smoothed_files)) {
  csv_path <- smoothed_files[i]
  session_name <- tools::file_path_sans_ext(basename(csv_path))
  toml_path <- file.path(toml_dir, paste0(session_name, ".toml"))
  
  roi_coords <- read_toml_roi(toml_path)

  if (is.null(roi_coords)) {
    warning(sprintf("Skipping %s - no ROI found", session_name))
    next
  }
  
  track_df <- read_csv(csv_path, show_col_types = FALSE)
  
  # Get the pixels_per_cm for this session
  px_per_cm <- track_df$pixels_per_cm[1]
  
  if (is.na(px_per_cm)) {
    warning(sprintf("Skipping %s - no calibration data", session_name))
    next
  }
  
  # Convert smoothed coordinates back to pixels for spatial analysis
  track_df <- track_df |>
    mutate(
      x_px = x_smooth_ind1 * px_per_cm,
      y_px = y_smooth_ind1 * px_per_cm
    ) |>
    filter(!is.na(x_px) & !is.na(y_px))
  
  # Create sf objects
  # Arena polygon
  arena_polygon <- st_polygon(list(roi_coords))
  arena_sf <- st_sfc(arena_polygon)
  arena_boundary <- st_boundary(arena_sf)
  
  # Get the demonstrator edge using the smart function
  demonstrator_edge <- get_demonstrator_edge(roi_coords, arena_sf)
  demonstrator_edge_sf <- st_sfc(demonstrator_edge)
  
  # Extract coordinates for plotting
  demonstrator_edge_coords <- st_coordinates(demonstrator_edge)
  
  # Fish positions as points
  fish_points <- st_as_sf(track_df, coords = c("x_px", "y_px"))
  
  # Calculate distances
  # Distance to arena edge (boundary)
  dist_to_edge_px <- as.numeric(st_distance(fish_points, arena_boundary))
  
  # Distance to demonstrator edge
  dist_to_top_px <- as.numeric(st_distance(fish_points, demonstrator_edge_sf))
  
  # Convert distances to cm
  track_df <- track_df |>
    mutate(
      dist_to_edge_cm = dist_to_edge_px / px_per_cm,
      dist_to_top_cm = dist_to_top_px / px_per_cm
    )
  
  # Check if points are inside the arena
  points_inside <- st_within(fish_points, arena_sf, sparse = FALSE)[, 1]
  track_df$inside_arena <- points_inside
  
  # Store results
  spatial_results[[session_name]] <- track_df
  
  # Create diagnostic plot for each session
  spatial_plot <- ggplot() +
    geom_polygon(data = as.data.frame(roi_coords), aes(x = V1, y = V2), fill = "lightblue", alpha = 0.3, color = "black", linewidth = 1) +
    geom_segment(aes(x = demonstrator_edge_coords[1, "X"], y = demonstrator_edge_coords[1, "Y"], xend = demonstrator_edge_coords[2, "X"], yend = demonstrator_edge_coords[2, "Y"]), color = "red", linewidth = 2) +
    geom_point(data = as.data.frame(demonstrator_edge_coords), aes(x = X, y = Y), color = "red", size = 3) +
    geom_path(data = track_df, aes(x = x_px, y = y_px), alpha = 0.3, color = "gray40") +
    geom_point(data = track_df, aes(x = x_px, y = y_px, color = dist_to_top_cm), size = 0.5, alpha = 0.5) +
    scale_color_viridis_c(option = "plasma", name = "Distance to\ndemo. (cm)") +
    scale_y_reverse() +
    coord_fixed() +
    labs(
      title = paste("Spatial Analysis:", session_name),
      subtitle = sprintf("Points inside arena: %.1f%% | Red line = demonstrator edge", mean(points_inside) * 100),
      x = "X (pixels)", y = "Y (pixels)") +
    theme_minimal() +
    theme(legend.position = "right")
  
  ggsave(filename = file.path(spatial_output_dir, paste0(session_name, "_spatial.png")), plot = spatial_plot, width = 10, height = 8, dpi = 150)
  
  # progress update
  if (i %% 25 == 0) cat("Processed", i, "of", length(smoothed_files), "files\n")
}

cat("Spatial processing complete!\n")
cat("Diagnostic plots saved to:", spatial_output_dir, "\n")
cat("Total plots saved:", length(spatial_results), "\n")
```

## 4.3 Update smoothed CSVs with spatial metrics

```{r save_spatial_metrics, message=FALSE}
updated_count <- 0

for (session_name in names(spatial_results)) {
  track_df <- spatial_results[[session_name]]
  
  spatial_cols <- track_df |>
    select(time, dist_to_edge_cm, dist_to_top_cm, inside_arena)
  
  csv_path <- file.path(smoothed_track_dir, paste0(session_name, ".csv"))
  original_df <- read_csv(csv_path, show_col_types = FALSE)
  
  original_df <- original_df |>
    select(-any_of(c("dist_to_edge_cm", "dist_to_top_cm", "inside_arena")))
  
  updated_df <- original_df |>
    left_join(spatial_cols, by = "time")
  
  write_csv(updated_df, csv_path)
  updated_count <- updated_count + 1
  print(updated_count)
}

cat("Updated", updated_count, "CSV files with spatial metrics.\n")
```

# 5. Final dataframe and blinding

The response variables of interest

*activity (focal)*

-   latency to start moving (time until sliding window proportion > 0.5) == "latency_to_moving"

-   Recovery to baseline behavioural levels (Recovery time was defined as the earliest changepoint occurring after 300 seconds where (1) the similarity at the
changepoint exceeded a threshold of 0.8, (2) the mean similarity from that changepoint to the end of the trial exceeded 0.8, and (3) the mean similarity after the changepoint was greater than the mean similarity of the preceding segment, indicating a sustained return to baseline-like behaviour rather than a transient fluctuation. Individuals that did not meet these criteria were classified as not having recovered within the trial period.)


## 5.1 Example track

Here is an example of one track with the mentioned response variables calculated.

```{r helper_functions}
# ---------------------------------------------------------------------------
# Helper: Detect time at which a fish returns to baseline behaviour
# Uses velocity and prop_moving to build a similarity score, with changepoint
# detection (PELT, BIC) on the smoothed similarity time-series.
#
# Arguments
#   trial_temporal   – per-second temporal df for the TRIAL phase
#   baseline_temporal – per-second temporal df for the matching BASELINE phase
#   velocity_col     – name of the velocity column (character)
#   moving_col       – name of the binary behavioural-state column (character)
#   rolling_k        – window width (seconds) for rolling means
#   smooth_k         – additional smoothing window for similarity
#   min_time         – ignore changepoints before this many seconds
#   similarity_threshold – minimum post-changepoint similarity to qualify
#
# Returns a named list: recovery_time, similarity_before, similarity_after,
#   returned_to_baseline (logical), baseline_velocity_mean, baseline_velocity_sd,
#   baseline_prop_moving_mean, baseline_prop_moving_sd
# ---------------------------------------------------------------------------
detect_recovery_time <- function(trial_temporal,
                                  baseline_temporal,
                                  velocity_col = "velocity_cm_s",
                                  moving_col = "behaviour_ind1_numeric",
                                  rolling_k = 60,
                                  smooth_k = 30,
                                  min_time = 300,
                                  similarity_threshold = 0.8) {

  # ---- empty-result template ------------------------------------------------
  empty_result <- list(
    recovery_time = NA_real_,
    similarity_before = NA_real_,
    similarity_after = NA_real_,
    returned_to_baseline = FALSE,
    baseline_velocity_mean = NA_real_,
    baseline_velocity_sd = NA_real_,
    baseline_prop_moving_mean = NA_real_,
    baseline_prop_moving_sd = NA_real_
  )

  # ---- guard: need baseline data --------------------------------------------
  if (is.null(baseline_temporal) || nrow(baseline_temporal) < rolling_k) {
    return(empty_result)
  }

  # ---- baseline profile -----------------------------------------------------
  bl_vel <- baseline_temporal[[velocity_col]]
  bl_move <- baseline_temporal[[moving_col]]

  bl_profile <- list(
    vel_mean = mean(bl_vel, na.rm = TRUE),
    vel_sd = sd(bl_vel, na.rm = TRUE),
    move_mean = mean(bl_move, na.rm = TRUE),
    move_sd = sd(bl_move, na.rm = TRUE)
  )

  # ---- trial rolling means --------------------------------------------------
  trial_vel <- trial_temporal[[velocity_col]]
  trial_move <- trial_temporal[[moving_col]]
  trial_time <- trial_temporal$time

  if (length(trial_vel) < rolling_k) return(empty_result)

  vel_rolling <- zoo::rollmean(trial_vel, k = rolling_k, fill = NA, align = "center")
  move_rolling <- zoo::rollmean(trial_move, k = rolling_k, fill = NA, align = "center")

  # ---- normalised deviations ------------------------------------------------
  # Only penalize when BELOW baseline (slower/less moving = still frozen)
  # Being above baseline (faster/more moving) is fine - not penalized
  vel_dev <- pmax(0, bl_profile$vel_mean - vel_rolling) /
    (bl_profile$vel_sd + 0.001)                        # velocity z-score distance (only if below)

  move_dev <- pmax(0, bl_profile$move_mean - move_rolling) /
    (bl_profile$move_sd + 0.001)                       # prop_moving z-score distance (only if below)

  similarity <- 1 / (1 + 0.5 * vel_dev + 0.5 * move_dev)
  sim_smooth <- zoo::rollmean(similarity, k = smooth_k, fill = NA, align = "center")

  # ---- changepoint detection ------------------------------------------------
  valid_idx <- which(!is.na(sim_smooth))
  if (length(valid_idx) < 100) return(empty_result)

  sim_valid <- sim_smooth[valid_idx]

  cpt <- tryCatch(
    changepoint::cpt.meanvar(
      sim_valid,
      method = "PELT",
      penalty = "BIC",
      minseglen = 60
    ),
    error = function(e) NULL
  )

  if (is.null(cpt)) return(empty_result)

  cps <- changepoint::cpts(cpt)
  seg_means <- changepoint::param.est(cpt)$mean

  # ---- find first qualifying recovery changepoint ---------------------------
  recovery_time <- NA_real_
  sim_before <- NA_real_
  sim_after <- NA_real_

  if (length(cps) > 0) {
    for (i in seq_along(cps)) {
      cp_idx  <- cps[i]
      cp_time <- trial_time[valid_idx[cp_idx]]
      if (cp_time < min_time) next
      
      # Similarity at the changepoint itself
      sim_at_cp <- sim_valid[cp_idx]
      
      # Calculate mean similarity from this changepoint to END of trial
      mean_after_cp <- mean(sim_valid[cp_idx:length(sim_valid)], na.rm = TRUE)
      
      if (i < length(seg_means)) {
        before <- seg_means[i]
        # Check: (1) changepoint itself >= threshold, (2) mean onwards >= threshold, (3) represents an increase
        if (sim_at_cp >= similarity_threshold && 
            mean_after_cp >= similarity_threshold && 
            mean_after_cp > before) {
          recovery_time <- cp_time
          sim_before <- before
          sim_after <- mean_after_cp
          break
        }
      }
    }
  }

  list(
    recovery_time = recovery_time,
    similarity_before = sim_before,
    similarity_after = sim_after,
    returned_to_baseline = !is.na(recovery_time),
    baseline_velocity_mean = bl_profile$vel_mean,
    baseline_velocity_sd = bl_profile$vel_sd,
    baseline_prop_moving_mean = bl_profile$move_mean,
    baseline_prop_moving_sd = bl_profile$move_sd
  )
}
```


```{r example_track_setup}
example_df <- read_csv("./2-data/3-behaviour_tracks/d1-t2-c1-t-1h2-i-focal.csv", show_col_types = FALSE)
```

### 5.1.1 Create per-second temporal dataframe

```{r example_temporal_df}
# Parameters
thigmotaxis_threshold_cm <- 3
proximity_threshold_cm <- 3
sliding_window_seconds <- 60

# Determine the time range where the FOCAL fish (ind1) has valid tracking data
# This ensures the temporal_df only covers the period when the focal fish is present
# (demonstrators may start earlier in trial sessions)
example_df_with_valid <- example_df |>
  mutate(
    has_valid_data = !is.na(x_smooth_ind1) & !is.na(y_smooth_ind1)
  )

valid_indices <- which(example_df_with_valid$has_valid_data)
first_valid_time <- example_df_with_valid$time[min(valid_indices)]
last_valid_time <- example_df_with_valid$time[max(valid_indices)]

cat("First valid tracking time:", first_valid_time, "s\n")
cat("Last valid tracking time:", last_valid_time, "s\n")

# Filter example_df to only include data within the valid time range
example_df_filtered <- example_df |>
  filter(time >= first_valid_time & time <= last_valid_time)

# Aggregate to per-second values
temporal_df <- example_df_filtered |>
  mutate(second = floor(time)) |>
  group_by(second) |>
  summarise(
    # Position (mean per second)
    x_cm = mean(x_smooth_ind1, na.rm = TRUE),
    y_cm = mean(y_smooth_ind1, na.rm = TRUE),
    
    # Velocity (mean per second)
    velocity_cm_s = mean(velocity_ind1, na.rm = TRUE),
    
    # behavioural state - proportion of frames moving in this second
    n_frames = n(),
    n_moving = sum(behaviour_ind1 == "moving", na.rm = TRUE),
    prop_moving_instant = n_moving / n_frames,
    
    # Demonstrator behavioural states - proportion of frames moving
    n_moving_ind2 = if ("behaviour_ind2" %in% names(example_df)) sum(behaviour_ind2 == "moving", na.rm = TRUE) else NA_real_,
    prop_moving_instant_ind2 = if ("behaviour_ind2" %in% names(example_df)) n_moving_ind2 / n_frames else NA_real_,
    
    n_moving_ind3 = if ("behaviour_ind3" %in% names(example_df)) sum(behaviour_ind3 == "moving", na.rm = TRUE) else NA_real_,
    prop_moving_instant_ind3 = if ("behaviour_ind3" %in% names(example_df)) n_moving_ind3 / n_frames else NA_real_,
    
    n_moving_ind4 = if ("behaviour_ind4" %in% names(example_df)) sum(behaviour_ind4 == "moving", na.rm = TRUE) else NA_real_,
    prop_moving_instant_ind4 = if ("behaviour_ind4" %in% names(example_df)) n_moving_ind4 / n_frames else NA_real_,
    
    # Spatial metrics (mean per second)
    dist_to_edge_cm = mean(dist_to_edge_cm, na.rm = TRUE),
    dist_to_top_cm = mean(dist_to_top_cm, na.rm = TRUE),
    
    # Binary thigmotaxis (proportion of frames within threshold)
    prop_near_edge = mean(dist_to_edge_cm < thigmotaxis_threshold_cm, na.rm = TRUE),
    
    # Binary social proximity (proportion of frames within threshold)
    prop_near_demonstrator = mean(dist_to_top_cm < proximity_threshold_cm, na.rm = TRUE),
    
    # Behavioural states as numeric (moving = 1, freezing = 0) 
    behaviour_ind1_numeric = mean(behaviour_ind1 == "moving", na.rm = TRUE),
    behaviour_ind2_numeric = if ("behaviour_ind2" %in% names(example_df)) mean(behaviour_ind2 == "moving", na.rm = TRUE) else NA_real_,
    behaviour_ind3_numeric = if ("behaviour_ind3" %in% names(example_df)) mean(behaviour_ind3 == "moving", na.rm = TRUE) else NA_real_,
    behaviour_ind4_numeric = if ("behaviour_ind4" %in% names(example_df)) mean(behaviour_ind4 == "moving", na.rm = TRUE) else NA_real_,
    
    .groups = "drop"
  ) |>
  rename(time = second) |>
  arrange(time)

# Add trial_time column (starts at 0 from first valid tracking data)
first_time_in_temporal <- min(temporal_df$time)
temporal_df <- temporal_df |>
  mutate(trial_time = time - first_time_in_temporal)

# Calculate sliding window proportion of time moving (1-minute window centered on each second)
# Using zoo::rollapply for the sliding window with partial=TRUE to handle edges
temporal_df <- temporal_df |>
  mutate(
    # Sliding window proportion moving (centered, 60s window)
    # partial=TRUE allows smaller windows at the edges
    prop_moving_sliding = zoo::rollapply(
      prop_moving_instant,
      width = sliding_window_seconds,
      FUN = mean,
      fill = NA,
      align = "center",
      na.rm = TRUE,
      partial = TRUE
    ),
    
    # Demonstrator sliding window proportions
    prop_moving_sliding_ind2 = zoo::rollapply(
      prop_moving_instant_ind2,
      width = sliding_window_seconds,
      FUN = mean,
      fill = NA,
      align = "center",
      na.rm = TRUE,
      partial = TRUE
    ),
    
    prop_moving_sliding_ind3 = zoo::rollapply(
      prop_moving_instant_ind3,
      width = sliding_window_seconds,
      FUN = mean,
      fill = NA,
      align = "center",
      na.rm = TRUE,
      partial = TRUE
    ),
    
    prop_moving_sliding_ind4 = zoo::rollapply(
      prop_moving_instant_ind4,
      width = sliding_window_seconds,
      FUN = mean,
      fill = NA,
      align = "center",
      na.rm = TRUE,
      partial = TRUE
    ),
    
    # Average demonstrator sliding window proportion
    prop_moving_sliding_demonstrators = rowMeans(
      cbind(prop_moving_sliding_ind2, prop_moving_sliding_ind3, prop_moving_sliding_ind4),
      na.rm = TRUE
    )
  )

# Add metadata columns
temporal_df <- temporal_df |>
  mutate(
    session_name = "d2-t1-c10-t-2l3-i",
    day = example_df$day[1],
    trial = example_df$trial[1],
    camera = example_df$camera[1],
    phase = example_df$phase[1],
    ID = example_df$ID[1],
    treatment = example_df$treatment[1],
    information = example_df$information[1]
  )

cat("Temporal dataframe created with", nrow(temporal_df), "rows (seconds)\n")
head(temporal_df)
```

### 5.1.2 Create whole-trial summary dataframe

```{r example_summary_df}
# Calculate whole-trial summary metrics
summary_df <- temporal_df |>
  summarise(
    # Metadata
    session_name = first(session_name),
    day = first(day),
    trial = first(trial),
    camera = first(camera),
    phase = first(phase),
    ID = first(ID),
    treatment = first(treatment),
    information = first(information),
    
    # Trial duration
    duration_s = max(time) - min(time) + 1,
    
    # ===== FREEZING METRICS =====
    # Total time spent moving (seconds where majority of frames were moving)
    total_time_moving = sum(prop_moving_instant > 0.5, na.rm = TRUE),
    
    # Proportion of trial spent moving
    prop_time_moving = total_time_moving / duration_s,
    
    # Latency to start moving more than freezing
    # (time until sliding window proportion > 0.5)
    latency_to_moving = {
      idx <- which(prop_moving_sliding > 0.5)
      if (length(idx) > 0) trial_time[min(idx)] else NA_real_
    },
    
    # Mean velocity
    mean_velocity_cm = mean(velocity_cm_s, na.rm = TRUE),
    
    # ===== REACTION TIME (time to first freeze) =====
    # Time until focal fish first freezes (sliding window proportion <= 0.5)
    reaction_time_ind1 = {
      idx <- which(behaviour_ind1_numeric <= 0.5)
      if (length(idx) > 0) trial_time[min(idx)] else NA_real_
    },
    # Distance from demonstrators at time of first freeze
    reaction_distance_from_dems = {
      idx <- which(behaviour_ind1_numeric <= 0.5)
      if (length(idx) > 0) dist_to_top_cm[min(idx)] else NA_real_
    },
    # Reaction time for each demonstrator
    reaction_time_ind2 = {
      idx <- which(behaviour_ind2_numeric <= 0.5)
      if (length(idx) > 0) trial_time[min(idx)] else NA_real_
    },
    reaction_time_ind3 = {
      idx <- which(behaviour_ind3_numeric <= 0.5)
      if (length(idx) > 0) trial_time[min(idx)] else NA_real_
    },
    reaction_time_ind4 = {
      idx <- which(behaviour_ind4_numeric <= 0.5)
      if (length(idx) > 0) trial_time[min(idx)] else NA_real_
    },
    # Average reaction time across demonstrators
    reaction_time_demonstrators = mean(
      c(reaction_time_ind2, reaction_time_ind3, reaction_time_ind4),
      na.rm = TRUE
    ),
    
    # ===== DEMONSTRATOR FREEZING METRICS =====
    # Proportion of time moving for each demonstrator
    prop_time_moving_ind2 = mean(prop_moving_instant_ind2 > 0.5, na.rm = TRUE),
    prop_time_moving_ind3 = mean(prop_moving_instant_ind3 > 0.5, na.rm = TRUE),
    prop_time_moving_ind4 = mean(prop_moving_instant_ind4 > 0.5, na.rm = TRUE),
    
    # Average proportion of time moving across demonstrators
    prop_time_moving_demonstrators = mean(c(prop_time_moving_ind2, prop_time_moving_ind3, prop_time_moving_ind4), na.rm = TRUE),
    
    # Latency to moving for each demonstrator
    latency_to_moving_ind2 = {
      idx <- which(prop_moving_sliding_ind2 > 0.5)
      if (length(idx) > 0) trial_time[min(idx)] else NA_real_
    },
    latency_to_moving_ind3 = {
      idx <- which(prop_moving_sliding_ind3 > 0.5)
      if (length(idx) > 0) trial_time[min(idx)] else NA_real_
    },
    latency_to_moving_ind4 = {
      idx <- which(prop_moving_sliding_ind4 > 0.5)
      if (length(idx) > 0) trial_time[min(idx)] else NA_real_
    },
    
    # Average latency to moving across demonstrators
    latency_to_moving_demonstrators = mean(c(latency_to_moving_ind2, latency_to_moving_ind3, latency_to_moving_ind4), na.rm = TRUE),
    
    # ===== THIGMOTAXIS METRICS =====
    # Total time within 3cm of edge
    total_time_near_edge = sum(prop_near_edge > 0.5, na.rm = TRUE),
    
    # Proportion of trial near edge
    prop_time_near_edge = total_time_near_edge / duration_s,
    
    # Mean distance to edge
    mean_dist_to_edge_cm = mean(dist_to_edge_cm, na.rm = TRUE),
    
    # ===== SOCIAL PROXIMITY METRICS =====
    # Total time within 3cm of demonstrator side (top edge)
    total_time_near_demonstrator = sum(prop_near_demonstrator > 0.5, na.rm = TRUE),
    
    # Proportion of trial near demonstrator
    prop_time_near_demonstrator = total_time_near_demonstrator / duration_s,
    
    # Mean distance to demonstrator side
    mean_dist_to_top_cm = mean(dist_to_top_cm, na.rm = TRUE)
  )

# ===== RECOVERY TIME (CHANGEPOINT DETECTION) =====
# Read the matching baseline file
baseline_path <- "./2-data/3-behaviour_tracks/d2-t1-c10-b-2l3-i-focal.csv"

baseline_temporal <- NULL
if (file.exists(baseline_path)) {
  bl_raw <- read_csv(baseline_path, show_col_types = FALSE) |>
    filter(between(time, 300, 1500))
  
  if (nrow(bl_raw) > 0) {
    baseline_temporal <- bl_raw |>
      mutate(second = floor(time)) |>
      group_by(second) |>
      summarise(
        velocity_cm_s = mean(velocity_ind1, na.rm = TRUE),
        behaviour_ind1_numeric = mean(behaviour_ind1 == "moving", na.rm = TRUE),
        .groups = "drop"
      ) |>
      rename(time = second)
  }
}

# Detect recovery time for focal fish
focal_recovery <- detect_recovery_time(
  trial_temporal    = temporal_df,
  baseline_temporal = baseline_temporal,
  velocity_col      = "velocity_cm_s",
  moving_col        = "behaviour_ind1_numeric"
)

# Add recovery metrics to summary
summary_df <- summary_df |>
  mutate(
    recovery_time = focal_recovery$recovery_time,
    returned_to_baseline = focal_recovery$returned_to_baseline,
    recovery_similarity_before = focal_recovery$similarity_before,
    recovery_similarity_after = focal_recovery$similarity_after,
    baseline_velocity_mean = focal_recovery$baseline_velocity_mean,
    baseline_velocity_sd = focal_recovery$baseline_velocity_sd,
    baseline_prop_moving_mean = focal_recovery$baseline_prop_moving_mean,
    baseline_prop_moving_sd = focal_recovery$baseline_prop_moving_sd,
    
    # ===== PRE-RECOVERY SPATIAL METRICS =====
    # Seconds near edge before recovery (or full trial if no recovery)
    time_near_edge_pre_recovery = {
      cutoff <- ifelse(is.na(focal_recovery$recovery_time),
                       max(temporal_df$time), focal_recovery$recovery_time)
      sum(temporal_df$prop_near_edge[temporal_df$time <= cutoff] > 0.5, na.rm = TRUE)
    },
    # Seconds near demonstrator before recovery (or full trial if no recovery)
    time_near_demonstrator_pre_recovery = {
      cutoff <- ifelse(is.na(focal_recovery$recovery_time),
                       max(temporal_df$time), focal_recovery$recovery_time)
      sum(temporal_df$prop_near_demonstrator[temporal_df$time <= cutoff] > 0.5, na.rm = TRUE)
    },
    # Seconds spent moving before recovery (or full trial if no recovery)
    time_moving_pre_recovery = {
      cutoff <- ifelse(is.na(focal_recovery$recovery_time),
                       max(temporal_df$time), focal_recovery$recovery_time)
      sum(temporal_df$prop_moving_instant[temporal_df$time <= cutoff] > 0.5, na.rm = TRUE)
    },
    # Mean distance to edge before recovery
    mean_dist_to_edge_pre_recovery = {
      cutoff <- ifelse(is.na(focal_recovery$recovery_time),
                       max(temporal_df$time), focal_recovery$recovery_time)
      mean(temporal_df$dist_to_edge_cm[temporal_df$time <= cutoff], na.rm = TRUE)
    },
    # Mean distance to demonstrator before recovery
    mean_dist_to_demonstrator_pre_recovery = {
      cutoff <- ifelse(is.na(focal_recovery$recovery_time),
                       max(temporal_df$time), focal_recovery$recovery_time)
      mean(temporal_df$dist_to_top_cm[temporal_df$time <= cutoff], na.rm = TRUE)
    }
  )

head(summary_df)
```

### 5.1.3 Visualize the example track

```{r example_visualization, fig.width=10, fig.height=6}
# Plot 1: Sliding window proportion moving over time (focal and demonstrators)
p1 <- ggplot(temporal_df, aes(x = trial_time)) +
  #geom_line(aes(y = prop_moving_sliding_demonstrators, color = "Demonstrators (avg)"), linewidth = 0.8) +
  geom_line(aes(y = prop_moving_sliding, color = "Focal"), linewidth = 0.8) +
  geom_hline(yintercept = 0.5, linetype = "dotted", color = "gray50", linewidth = 0.3) +
  geom_hline(yintercept = 1, color = "gray30", linewidth = 0.3) +
  geom_vline(
    xintercept = summary_df$latency_to_moving,
    linetype = "dotted", color = "darkgreen", linewidth = 0.8
  ) +
  annotate(
    "text",
    x = summary_df$latency_to_moving - 930,
    y = 1.1,
    label = paste0("Latency to move: ", round(summary_df$latency_to_moving, 0), "s"),
    color = "darkgreen", hjust = 0, size = 3
  ) +
  # Recovery time vline
  {if (!is.na(summary_df$recovery_time))
    geom_vline(
      xintercept = summary_df$recovery_time,
      linetype = "dashed", color = "red", linewidth = 0.8
    )} +
  {if (!is.na(summary_df$recovery_time))
    annotate(
      "text",
      x = summary_df$recovery_time + 50,
      y = 1.1,
      label = paste("Recovery:", round(summary_df$recovery_time, 0), "s"),
      color = "red", hjust = 0, size = 3
    )} +
  scale_color_manual(
    name = "Individual",
    values = c("Focal" = "#1F78B4", "Demonstrators (avg)" = "#E31A1C")
  ) +
  scale_y_continuous(limits = c(0, 1.1), breaks = c(0, 0.2, 0.4, 0.6, 0.8, 1)) +
  labs(
    title = "Proportion of Time Moving (60s sliding window)",
    x = NULL, y = "Proportion Moving"
  ) +
  theme_classic() +
  theme(
    axis.text.x = element_blank(),
    legend.position = "none"
  )

# Plot 2: Velocity over time
p2 <- ggplot(temporal_df, aes(x = trial_time, y = velocity_cm_s)) +
  geom_line(color = "gray40", alpha = 0.7) +
  geom_smooth(method = "loess", span = 0.1, se = FALSE, color = "#1F78B4") +
  labs(
    title = "Velocity Over Time",
    x = NULL, y = "Velocity (cm/s)"
  ) +
  theme_classic() +
  theme(axis.text.x = element_blank())

# Plot 3: Distance to edge over time
p3 <- ggplot(temporal_df, aes(x = trial_time, y = dist_to_edge_cm)) +
  geom_line(color = "gray40", alpha = 0.7) +
  geom_hline(yintercept = 3, colour = "orange", linetype = "dashed") +
  geom_smooth(method = "loess", span = 0.1, se = FALSE, color = "#FF7F00") +
  labs(
    title = "Distance to Arena Edge (Thigmotaxis)",
    x = NULL, y = "Distance (cm)"
  ) +
  theme_classic() +
  theme(axis.text.x = element_blank())

# Plot 4: Distance to demonstrator side over time
p4 <- ggplot(temporal_df, aes(x = trial_time, y = dist_to_top_cm)) +
  geom_line(color = "gray40", alpha = 0.7) +
  geom_hline(yintercept = 3, colour = "purple", linetype = "dashed") +
  geom_smooth(method = "loess", span = 0.1, se = FALSE, color = "#6A3D9A") +
  labs(
    title = "Distance to Demonstrator Side (Social Proximity)",
    x = "Time (s)", y = "Distance (cm)"
  ) +
  theme_classic()

# Combine plots
combined_plot <- (p1 / p2 / p3 / p4 +
  plot_annotation(
    subtitle = sprintf(
      "Focal moving: %.1f%% | Demo moving: %.1f%% | Near edge: %.1f%% | Near demonstrator: %.1f%% | Recovery: %s",
      summary_df$prop_time_moving * 100,
      summary_df$prop_time_moving_demonstrators * 100,
      summary_df$prop_time_near_edge * 100,
      summary_df$prop_time_near_demonstrator * 100,
      ifelse(is.na(summary_df$recovery_time), "none",
             paste0(round(summary_df$recovery_time), "s"))
    )
  ))

combined_plot 

ggsave("./3-figs/combined_plot.png", combined_plot, height = 6, width = 10, dpi = 300)
```



## 5.2 Batch process all smoothed tracks

```{r batch_process_all_tracks}
# Parameters
thigmotaxis_threshold_cm <- 3
proximity_threshold_cm <- 3
sliding_window_seconds <- 60

# Get all behaviour track files (with HMM-decoded behavioural states)
smoothed_track_dir <- "./2-data/3-behaviour_tracks"

smoothed_files <- list.files(
  path = smoothed_track_dir,
  pattern = "\\.csv$",
  full.names = TRUE) |>
  (\(x) x[!grepl("^\\._", basename(x))])()

cat(sprintf("Found %d behaviour track files\n", length(smoothed_files)))

# Function to process a single track
process_track <- function(csv_path) {
  session_name <- tools::file_path_sans_ext(basename(csv_path))
  
  # Read the track
  track_df <- tryCatch(
    read_csv(csv_path, show_col_types = FALSE),
    error = function(e) {
      warning(sprintf("Failed to read %s: %s", session_name, e$message))
      return(NULL)
    }
  )
  
  if (is.null(track_df)) return(NULL)

  # Check if phase column exists
  if (!"phase" %in% names(track_df)) {
    warning(sprintf("No phase column in %s", session_name))
    return(NULL)
  }
  
  # Check if required columns exist
  required_cols <- c("time", "x_smooth_ind1", "y_smooth_ind1", "velocity_ind1", 
                     "behaviour_ind1", "dist_to_edge_cm", "dist_to_top_cm")
  if (!all(required_cols %in% names(track_df))) {
    warning(sprintf("Missing required columns in %s", session_name))
    return(NULL)
  }
  
  # Determine the time range where tracking data exists for ANY individual
  # Determine the time range where the FOCAL fish (ind1) has valid tracking data
  # This ensures the temporal_df only covers the period when the focal fish is present
  # (demonstrators may start earlier in trial sessions)
  track_df_with_valid <- track_df |>
    mutate(
      has_valid_data = !is.na(x_smooth_ind1) & !is.na(y_smooth_ind1)
    )
  
  valid_indices <- which(track_df_with_valid$has_valid_data)
  
  if (length(valid_indices) == 0) {
    warning(sprintf("No valid tracking data found in %s", session_name))
    return(NULL)
  }
  
  first_valid_time <- track_df_with_valid$time[min(valid_indices)]
  last_valid_time <- track_df_with_valid$time[max(valid_indices)]
  
  # Filter track_df to only include data within the valid time range
  track_df_filtered <- track_df |>
    filter(time >= first_valid_time & time <= last_valid_time)
  
  # Aggregate to per-second values
  temporal_df <- track_df_filtered |>
    mutate(second = floor(time)) |>
    group_by(second) |>
    summarise(
      # Position (mean per second)
      x_cm = mean(x_smooth_ind1, na.rm = TRUE),
      y_cm = mean(y_smooth_ind1, na.rm = TRUE),
      
      # Velocity (mean per second)
      velocity_cm_s = mean(velocity_ind1, na.rm = TRUE),
      
      # behavioural state - proportion of frames moving in this second
      n_frames = n(),
      n_moving = sum(behaviour_ind1 == "moving", na.rm = TRUE),
      prop_moving_instant = n_moving / n_frames,
      
      # Demonstrator behavioural states - proportion of frames moving
      n_moving_ind2 = if ("behaviour_ind2" %in% names(track_df)) sum(behaviour_ind2 == "moving", na.rm = TRUE) else NA_real_,
      prop_moving_instant_ind2 = if ("behaviour_ind2" %in% names(track_df)) n_moving_ind2 / n_frames else NA_real_,
      
      n_moving_ind3 = if ("behaviour_ind3" %in% names(track_df)) sum(behaviour_ind3 == "moving", na.rm = TRUE) else NA_real_,
      prop_moving_instant_ind3 = if ("behaviour_ind3" %in% names(track_df)) n_moving_ind3 / n_frames else NA_real_,
      
      n_moving_ind4 = if ("behaviour_ind4" %in% names(track_df)) sum(behaviour_ind4 == "moving", na.rm = TRUE) else NA_real_,
      prop_moving_instant_ind4 = if ("behaviour_ind4" %in% names(track_df)) n_moving_ind4 / n_frames else NA_real_,
      
      # Spatial metrics (mean per second)
      dist_to_edge_cm = mean(dist_to_edge_cm, na.rm = TRUE),
      dist_to_top_cm = mean(dist_to_top_cm, na.rm = TRUE),
      
      # Binary thigmotaxis (proportion of frames within threshold)
      prop_near_edge = mean(dist_to_edge_cm < thigmotaxis_threshold_cm, na.rm = TRUE),
      
      # Binary social proximity (proportion of frames within threshold)
      prop_near_demonstrator = mean(dist_to_top_cm < proximity_threshold_cm, na.rm = TRUE),
      
      # behavioural states for demonstrators
      # Convert to numeric: moving = 1, freezing = 0
      behaviour_ind1_numeric = mean(behaviour_ind1 == "moving", na.rm = TRUE),
      behaviour_ind2_numeric = if ("behaviour_ind2" %in% names(track_df)) mean(behaviour_ind2 == "moving", na.rm = TRUE) else NA_real_,
      behaviour_ind3_numeric = if ("behaviour_ind3" %in% names(track_df)) mean(behaviour_ind3 == "moving", na.rm = TRUE) else NA_real_,
      behaviour_ind4_numeric = if ("behaviour_ind4" %in% names(track_df)) mean(behaviour_ind4 == "moving", na.rm = TRUE) else NA_real_,
      
      .groups = "drop"
    ) |>
    rename(time = second) |>
    arrange(time)
  
  # Add trial_time column (starts at 0 from first valid tracking data)
  first_time_in_temporal <- min(temporal_df$time)
  temporal_df <- temporal_df |>
    mutate(trial_time = time - first_time_in_temporal)
  
  # Calculate sliding window proportion of time moving
  temporal_df <- temporal_df |>
    mutate(
      prop_moving_sliding = zoo::rollapply(
        prop_moving_instant,
        width = sliding_window_seconds,
        FUN = mean,
        fill = NA,
        align = "center",
        na.rm = TRUE,
        partial = TRUE
      ),
      
      # Demonstrator sliding window proportions
      prop_moving_sliding_ind2 = zoo::rollapply(
        prop_moving_instant_ind2,
        width = sliding_window_seconds,
        FUN = mean,
        fill = NA,
        align = "center",
        na.rm = TRUE,
        partial = TRUE
      ),
      
      prop_moving_sliding_ind3 = zoo::rollapply(
        prop_moving_instant_ind3,
        width = sliding_window_seconds,
        FUN = mean,
        fill = NA,
        align = "center",
        na.rm = TRUE,
        partial = TRUE
      ),
      
      prop_moving_sliding_ind4 = zoo::rollapply(
        prop_moving_instant_ind4,
        width = sliding_window_seconds,
        FUN = mean,
        fill = NA,
        align = "center",
        na.rm = TRUE,
        partial = TRUE
      ),
      
      # Average demonstrator sliding window proportion
      prop_moving_sliding_demonstrators = rowMeans(
        cbind(prop_moving_sliding_ind2, prop_moving_sliding_ind3, prop_moving_sliding_ind4),
        na.rm = TRUE
      )
    )
  
  # Add metadata columns
  temporal_df <- temporal_df |>
    mutate(
      session_name = session_name,
      day = track_df$day[1],
      trial = track_df$trial[1],
      camera = track_df$camera[1],
      phase = track_df$phase[1],
      ID = track_df$ID[1],
      treatment = track_df$treatment[1],
      information = track_df$information[1]
    )
  
  # Calculate whole-trial summary metrics
  summary_df <- temporal_df |>
    summarise(
      # Metadata
      session_name = first(session_name),
      day = first(day),
      trial = first(trial),
      camera = first(camera),
      phase = first(phase),
      ID = first(ID),
      treatment = first(treatment),
      information = first(information),
      
      # Trial duration
      duration_s = max(time) - min(time) + 1,
      
      # ===== FREEZING METRICS =====
      total_time_moving = sum(prop_moving_instant > 0.5, na.rm = TRUE),
      prop_time_moving = total_time_moving / duration_s,
      latency_to_moving = {
        idx <- which(prop_moving_sliding > 0.5)
        if (length(idx) > 0) trial_time[min(idx)] else NA_real_
      },
      mean_velocity_cm = mean(velocity_cm_s, na.rm = TRUE),
      
      # ===== REACTION TIME (time to first freeze) =====
      # Time until focal fish first freezes (sliding window proportion <= 0.5)
      reaction_time_ind1 = {
        idx <- which(behaviour_ind1_numeric <= 0.5)
        if (length(idx) > 0) trial_time[min(idx)] else NA_real_
      },
      # Distance from demonstrators at time of first freeze
      reaction_distance_from_dems = {
        idx <- which(behaviour_ind1_numeric <= 0.5)
        if (length(idx) > 0) dist_to_top_cm[min(idx)] else NA_real_
      },
      # Reaction time for each demonstrator
      reaction_time_ind2 = {
        idx <- which(behaviour_ind2_numeric <= 0.5)
        if (length(idx) > 0) trial_time[min(idx)] else NA_real_
      },
      reaction_time_ind3 = {
        idx <- which(behaviour_ind3_numeric <= 0.5)
        if (length(idx) > 0) trial_time[min(idx)] else NA_real_
      },
      reaction_time_ind4 = {
        idx <- which(behaviour_ind4_numeric <= 0.5)
        if (length(idx) > 0) trial_time[min(idx)] else NA_real_
      },
      # Average reaction time across demonstrators
      reaction_time_demonstrators = mean(
        c(reaction_time_ind2, reaction_time_ind3, reaction_time_ind4),
        na.rm = TRUE
      ),
      
      # ===== DEMONSTRATOR FREEZING METRICS =====
      # Proportion of time moving for each demonstrator
      prop_time_moving_ind2 = mean(prop_moving_instant_ind2 > 0.5, na.rm = TRUE),
      prop_time_moving_ind3 = mean(prop_moving_instant_ind3 > 0.5, na.rm = TRUE),
      prop_time_moving_ind4 = mean(prop_moving_instant_ind4 > 0.5, na.rm = TRUE),
      
      # Average proportion of time moving across demonstrators
      prop_time_moving_demonstrators = mean(c(prop_time_moving_ind2, prop_time_moving_ind3, prop_time_moving_ind4), na.rm = TRUE),
      
      # Latency to moving for each demonstrator
      latency_to_moving_ind2 = {
        idx <- which(prop_moving_sliding_ind2 > 0.5)
        if (length(idx) > 0) trial_time[min(idx)] else NA_real_
      },
      latency_to_moving_ind3 = {
        idx <- which(prop_moving_sliding_ind3 > 0.5)
        if (length(idx) > 0) trial_time[min(idx)] else NA_real_
      },
      latency_to_moving_ind4 = {
        idx <- which(prop_moving_sliding_ind4 > 0.5)
        if (length(idx) > 0) trial_time[min(idx)] else NA_real_
      },
      
      # Average latency to moving across demonstrators
      latency_to_moving_demonstrators = mean(c(latency_to_moving_ind2, latency_to_moving_ind3, latency_to_moving_ind4), na.rm = TRUE),
      
      # ===== THIGMOTAXIS METRICS =====
      total_time_near_edge = sum(prop_near_edge > 0.5, na.rm = TRUE),
      prop_time_near_edge = total_time_near_edge / duration_s,
      mean_dist_to_edge_cm = mean(dist_to_edge_cm, na.rm = TRUE),
      
      # ===== SOCIAL PROXIMITY METRICS =====
      total_time_near_demonstrator = sum(prop_near_demonstrator > 0.5, na.rm = TRUE),
      prop_time_near_demonstrator = total_time_near_demonstrator / duration_s,
      mean_dist_to_top_cm = mean(dist_to_top_cm, na.rm = TRUE)
    )
  
  # ===== RECOVERY TIME (CHANGEPOINT DETECTION) =====
  # Only meaningful for trial sessions – we compare against the matching baseline
  phase_label <- temporal_df$phase[1]
  
  if (!is.na(phase_label) && phase_label == "trial") {
    # Build the baseline file path (swap "-t-" for "-b-" in the session name)
    baseline_session <- stringr::str_replace(session_name, "-t-", "-b-")
    baseline_path <- file.path(dirname(csv_path), paste0(baseline_session, ".csv"))
    
    baseline_temporal <- NULL
    if (file.exists(baseline_path)) {
      bl_raw <- tryCatch(
        read_csv(baseline_path, show_col_types = FALSE),
        error = function(e) NULL
      )
      if (!is.null(bl_raw) && nrow(bl_raw) > 0) {
        # Trim baseline to the standardised 300-1500 s window
        bl_raw <- bl_raw |> filter(between(time, 300, 1500))
        
        # Aggregate baseline to per-second (mirrors the trial temporal_df)
        baseline_temporal <- bl_raw |>
          mutate(second = floor(time)) |>
          group_by(second) |>
          summarise(
            velocity_cm_s = mean(velocity_ind1, na.rm = TRUE),
            behaviour_ind1_numeric = mean(behaviour_ind1 == "moving", na.rm = TRUE),
            behaviour_ind2_numeric = if ("behaviour_ind2" %in% names(bl_raw)) mean(behaviour_ind2 == "moving", na.rm = TRUE) else NA_real_,
            behaviour_ind3_numeric = if ("behaviour_ind3" %in% names(bl_raw)) mean(behaviour_ind3 == "moving", na.rm = TRUE) else NA_real_,
            behaviour_ind4_numeric = if ("behaviour_ind4" %in% names(bl_raw)) mean(behaviour_ind4 == "moving", na.rm = TRUE) else NA_real_,
            .groups = "drop"
          ) |>
          rename(time = second)
      }
    }
    
    # --- Focal recovery ---
    focal_recovery <- detect_recovery_time(
      trial_temporal = temporal_df,
      baseline_temporal = baseline_temporal,
      velocity_col = "velocity_cm_s",
      moving_col = "behaviour_ind1_numeric"
    )
    
    # --- Demonstrator recovery (ind2, ind3, ind4) ---
    demo_recovery <- lapply(c("ind2", "ind3", "ind4"), function(ind) {
      vel_col <- "velocity_cm_s" # demonstrators share focal velocity proxy
      move_col <- paste0("behaviour_", ind, "_numeric")
      
      # Only attempt if the column exists in the trial temporal_df
      if (!move_col %in% names(temporal_df)) {
        return(list(recovery_time = NA_real_,
                    returned_to_baseline = FALSE,
                    baseline_prop_moving_sd = NA_real_))
      }
      
      # Build a mini baseline_temporal for this demonstrator
      bl_temp <- baseline_temporal
      if (!is.null(bl_temp) && !move_col %in% names(bl_temp)) bl_temp <- NULL
      
      detect_recovery_time(
        trial_temporal = temporal_df,
        baseline_temporal = bl_temp,
        velocity_col = vel_col,
        moving_col = move_col
      )
    })
    names(demo_recovery) <- c("ind2", "ind3", "ind4")
    
    # Append recovery columns to summary_df
    summary_df <- summary_df |>
      mutate(
        # ===== FOCAL RECOVERY =====
        recovery_time = focal_recovery$recovery_time,
        returned_to_baseline = focal_recovery$returned_to_baseline,
        recovery_similarity_before = focal_recovery$similarity_before,
        recovery_similarity_after = focal_recovery$similarity_after,
        baseline_velocity_mean = focal_recovery$baseline_velocity_mean,
        baseline_velocity_sd = focal_recovery$baseline_velocity_sd,
        baseline_prop_moving_mean = focal_recovery$baseline_prop_moving_mean,
        baseline_prop_moving_sd = focal_recovery$baseline_prop_moving_sd,
        
        # ===== DEMONSTRATOR RECOVERY =====
        recovery_time_ind2 = demo_recovery$ind2$recovery_time,
        returned_to_baseline_ind2 = demo_recovery$ind2$returned_to_baseline,
        baseline_prop_moving_sd_ind2 = demo_recovery$ind2$baseline_prop_moving_sd,
    
        recovery_time_ind3 = demo_recovery$ind3$recovery_time,
        returned_to_baseline_ind3 = demo_recovery$ind3$returned_to_baseline,
        baseline_prop_moving_sd_ind3 = demo_recovery$ind3$baseline_prop_moving_sd,
        
        recovery_time_ind4 = demo_recovery$ind4$recovery_time,
        returned_to_baseline_ind4 = demo_recovery$ind4$returned_to_baseline,
        baseline_prop_moving_sd_ind4 = demo_recovery$ind4$baseline_prop_moving_sd,
        
        # Average demonstrator recovery time
        recovery_time_demonstrators = mean(
          c(demo_recovery$ind2$recovery_time,
            demo_recovery$ind3$recovery_time,
            demo_recovery$ind4$recovery_time),
          na.rm = TRUE
        ),
        
        # ===== PRE-RECOVERY SPATIAL METRICS =====
        # Seconds near edge before recovery (or full trial if no recovery)
        time_near_edge_pre_recovery = {
          cutoff <- ifelse(is.na(focal_recovery$recovery_time),
                           max(temporal_df$time), focal_recovery$recovery_time)
          sum(temporal_df$prop_near_edge[temporal_df$time <= cutoff] > 0.5, na.rm = TRUE)
        },
        # Seconds near demonstrator before recovery (or full trial if no recovery)
        time_near_demonstrator_pre_recovery = {
          cutoff <- ifelse(is.na(focal_recovery$recovery_time),
                           max(temporal_df$time), focal_recovery$recovery_time)
          sum(temporal_df$prop_near_demonstrator[temporal_df$time <= cutoff] > 0.5, na.rm = TRUE)
        },
        # Seconds spent moving before recovery (or full trial if no recovery)
        time_moving_pre_recovery = {
          cutoff <- ifelse(is.na(focal_recovery$recovery_time),
                           max(temporal_df$time), focal_recovery$recovery_time)
          sum(temporal_df$prop_moving_instant[temporal_df$time <= cutoff] > 0.5, na.rm = TRUE)
        },
        # Mean distance to edge before recovery
        mean_dist_to_edge_pre_recovery = {
          cutoff <- ifelse(is.na(focal_recovery$recovery_time),
                           max(temporal_df$time), focal_recovery$recovery_time)
          mean(temporal_df$dist_to_edge_cm[temporal_df$time <= cutoff], na.rm = TRUE)
        },
        # Mean distance to demonstrator before recovery
        mean_dist_to_demonstrator_pre_recovery = {
          cutoff <- ifelse(is.na(focal_recovery$recovery_time),
                           max(temporal_df$time), focal_recovery$recovery_time)
          mean(temporal_df$dist_to_top_cm[temporal_df$time <= cutoff], na.rm = TRUE)
        }
      )
  } else {
    # Baseline sessions: fill recovery columns with NA
    summary_df <- summary_df |>
      mutate(
        recovery_time = NA_real_,
        returned_to_baseline = NA,
        recovery_similarity_before = NA_real_,
        recovery_similarity_after = NA_real_,
        baseline_velocity_mean = NA_real_,
        baseline_velocity_sd = NA_real_,
        baseline_prop_moving_mean = NA_real_,
        baseline_prop_moving_sd = NA_real_,
        recovery_time_ind2 = NA_real_,
        returned_to_baseline_ind2 = NA,
        baseline_prop_moving_sd_ind2 = NA_real_,
        recovery_time_ind3 = NA_real_,
        returned_to_baseline_ind3 = NA,
        baseline_prop_moving_sd_ind3 = NA_real_,
        recovery_time_ind4 = NA_real_,
        returned_to_baseline_ind4 = NA,
        baseline_prop_moving_sd_ind4 = NA_real_,
        recovery_time_demonstrators = NA_real_,
        time_near_edge_pre_recovery = NA_real_,
        time_near_demonstrator_pre_recovery = NA_real_,
        time_moving_pre_recovery = NA_real_,
        mean_dist_to_edge_pre_recovery = NA_real_,
        mean_dist_to_demonstrator_pre_recovery = NA_real_
      )
  }
  
  return(list(temporal = temporal_df, summary = summary_df))
}

# Set up parallel processing
plan(multisession, workers = max(1, availableCores() - 1))

cat(sprintf("Processing tracks using %d CPU cores...\n", nbrOfWorkers()))

# Process all tracks in parallel
results <- future_map(
  smoothed_files,
  process_track,
  .options = furrr_options(seed = TRUE),
  .progress = TRUE
)

# Reset to sequential processing
plan(sequential)

# Filter out NULL results and separate temporal and summary data
results_valid <- results[!sapply(results, is.null)]

cat(sprintf("\nSuccessfully processed %d tracks out of %d total files\n", 
            length(results_valid), length(smoothed_files)))

# Combine all temporal dataframes
temporal_df <- map_dfr(results_valid, ~.x$temporal)

# Combine all summary dataframes
summary_df <- map_dfr(results_valid, ~.x$summary)
```

## 5.3 Attached morphs

```{r}
morphs_raw <- read_csv("./2-data/morphs.csv") |> 
  rename(demonstrator_id = demonstrator) |> 
  select(session_name, weight, demonstrator_id)

# The morphs file only contains trial-phase session names (with "-t-")
# We need to also create baseline versions (with "-b-") since the weight 
# and demonstrator_id are the same for both phases of the same fish
morphs_trial <- morphs_raw

morphs_baseline <- morphs_raw |>
  mutate(session_name = str_replace(session_name, "-t-", "-b-"))

# Combine trial and baseline morphs
morphs <- bind_rows(morphs_trial, morphs_baseline)

cat(sprintf("Morphs data expanded from %d trial sessions to %d total (trial + baseline)\n",
            nrow(morphs_trial), nrow(morphs)))

all_temporal_morph_df <- merge(temporal_df, morphs, by ="session_name") |> 
  mutate(time = time/60,
         trial_time = trial_time/60)

all_summary_morph_df <- merge(summary_df, morphs, by ="session_name") |> 
  mutate(duration_m = duration_s/60,
         across(c(total_time_moving, starts_with("latency"), starts_with("recovery_time"),
                  starts_with("time_"), total_time_near_edge, total_time_near_demonstrator),
                ~ .x / 60))

cat(sprintf("Temporal data: %d rows (from %d before morph merge)\n", 
            nrow(all_temporal_morph_df), nrow(temporal_df)))
cat(sprintf("Summary data: %d rows (from %d before morph merge)\n", 
            nrow(all_summary_morph_df), nrow(summary_df)))
```


## 5.4 Create blinded versions of the data

This section creates blinded versions of both dataframes where treatment and information levels are randomized. A key file is saved that can be used to unblind the data after analysis.

```{r blind_data}
set.seed(9)

# Create random blinding codes for treatment levels
treatment_levels <- c("c", "l", "h")
treatment_blind_codes <- c("TX", "TY", "TZ") 
treatment_shuffled <- sample(treatment_blind_codes) 

# Create random blinding codes for information levels
information_levels <- c("personal", "social")
information_blind_codes <- c("IA", "IB")
information_shuffled <- sample(information_blind_codes)

# Create blinding key (mapping from real to blinded)
blinding_key <- tibble(
  variable = c(rep("treatment", 3), rep("information", 2)),
  original = c(treatment_levels, information_levels),
  blinded = c(treatment_shuffled, information_shuffled))

# Function to apply blinding to a dataframe
apply_blinding <- function(df, blinding_key) {
  df_blinded <- df
  
  # Remove "-focal" suffix from session_name and ID before blinding
  df_blinded$session_name <- gsub("-focal$", "", df_blinded$session_name)
  if ("ID" %in% names(df_blinded)) {
    df_blinded$ID <- gsub("-focal$", "", df_blinded$ID)
  }
  
  # Blind treatment
  treatment_map <- setNames(
    blinding_key$blinded[blinding_key$variable == "treatment"],
    blinding_key$original[blinding_key$variable == "treatment"]
  )
  df_blinded$treatment <- treatment_map[df_blinded$treatment]
  
  # Blind information
  information_map <- setNames(
    blinding_key$blinded[blinding_key$variable == "information"],
    blinding_key$original[blinding_key$variable == "information"]
  )
  df_blinded$information <- information_map[df_blinded$information]
  
  # Blind session_name and ID (replace original codes with blinded ones)
  for (i in 1:nrow(blinding_key)) {
    pattern <- blinding_key$original[i]
    replacement <- blinding_key$blinded[i]
    df_blinded$session_name <- gsub(pattern, replacement, df_blinded$session_name)
    if ("ID" %in% names(df_blinded)) {
      df_blinded$ID <- gsub(pattern, replacement, df_blinded$ID)
    }
  }
  
  return(df_blinded)
}

temporal_df_blinded <- apply_blinding(all_temporal_morph_df, blinding_key)
summary_df_blinded <- apply_blinding(all_summary_morph_df, blinding_key)

write_csv(temporal_df_blinded, "./2-data/temporal_data.csv")
write_csv(summary_df_blinded, "./2-data/summary_data.csv")
write_csv(blinding_key, "./2-data/BLINDING_KEY.csv")
```


